---
title: 'Kramann - Kabgani: PDGFRb+ cells - Kidney Model - RNA-Seq'
author:
- affiliation: Interdisciplinary Center for Clinical Research - RWTH-Aachen
  name: Ali T. Abdallah
  fontsize: 15pt
date: "01.06.2021"
output: 
  html_notebook:
    highlight: kate
    theme: paper
---

<style>
.main-container{
  min-width: 1200px;!important
  margin-left: auto;!important
  margin-right: auto;!important
  font-size: 15px;!important
}
.tabset{
  font-size: 15px;!important
}

table {
  width: 100%;
  border-collapse: collapse;
  border: 1px solid black;
}
tr.sub th {
    border-top-width: 1px;
    border-top-style: solid;
    border-top-color: #8ebffc;
    font-weight: normal;
    color: #333333;
}
.download{
	padding: 10px 15px;
	background: #4479BA;
	color: #FFF;
	-webkit-border-radius: 4px;
	-moz-border-radius: 4px;
	border-radius: 4px;
	border: solid 1px #20538D;
	text-shadow: 0 -1px 0 rgba(0, 0, 0, 0.4);
	-webkit-box-shadow: inset 0 1px 0 rgba(255, 255, 255, 0.4), 0 1px 1px rgba(0, 0, 0, 0.2);
	-moz-box-shadow: inset 0 1px 0 rgba(255, 255, 255, 0.4), 0 1px 1px rgba(0, 0, 0, 0.2);
	box-shadow: inset 0 1px 0 rgba(255, 255, 255, 0.4), 0 1px 1px rgba(0, 0, 0, 0.2);
	-webkit-transition-duration: 0.2s;
	-moz-transition-duration: 0.2s;
	transition-duration: 0.2s;
	-webkit-user-select:none;
	-moz-user-select:none;
	-ms-user-select:none;
	user-select:none;
}
.download:hover{
	background: #356094;
	border: solid 1px #2A4E77;
	text-decoration: none;
}
.download:active{
	-webkit-box-shadow: inset 0 1px 4px rgba(0, 0, 0, 0.6);
	-moz-box-shadow: inset 0 1px 4px rgba(0, 0, 0, 0.6);
	box-shadow: inset 0 1px 4px rgba(0, 0, 0, 0.6);
	background: #2E5481;
	border: solid 1px #203E5F;
}
h3{
color: #800000;
font-weight: bold;
}
h4{
color: #000080;
font-weight: bold;
}
h5{
color: #008080;
font-weight: bold
}
</style>


## {.tabset .tabset-fade }

```{r setup, echo=F}
library("RColorBrewer")
library(DESeq2)
library(ggplot2)
library(cowplot)
library(ComplexHeatmap)
concatenate <- function(Gene, Gene.Space=10, GeneClass=""){
   return(paste(Gene, paste(rep(" ",Gene.Space-nchar(Gene)),collapse=" "),"",GeneClass,sep=""))
}

```


<h3>Quality Control Plots</h3>

### ATAC-Seq
Here we very nicely see on both the PCA-plot and the Similarity heatmap precisely the expected behavior. Interestingly almost all effects of the treatment are explained by the PC1 (97% variance). On the other hand, the hierarchical clustering based on the euclidean distance between the samples' expression profiles reflects the expected behavior—a beneficial quality check.

```{r fig.width=13, fig.height=6, echo=F}
distances <- c("euclidean")
grids <- vector("list", length(distances))

for(d in 1:length(distances)){
    vsd <- vst(DDS.atac, blind = F)
    sampleDists <- dist(t(assay(vsd)), method = distances[d])
    sampleDistMatrix <- as.matrix(sampleDists)
    rownames(sampleDistMatrix) <- colnames(vsd)
    colnames(sampleDistMatrix) <- colnames(vsd)
    colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
    hm <- pheatmap::pheatmap(sampleDistMatrix,
                             clustering_distance_rows=sampleDists, 
                             clustering_distance_cols=sampleDists,
                             col=colors)
    hm <- ggplotify::as.ggplot(hm) 
}

pdf(file="Data/Images/QC.ATACSeq.pdf", width=13, height=6)
cowplot::plot_grid(ggPCA + xlab("PC1: 97% variance") + ylab("PC2: 1% variance"),
                   ggplot() + theme_void(), 
                   hm + theme(plot.title = element_text(hjust = 0.5, face="bold"),
                                                       plot.subtitle = element_text(hjust = 0.5)) + 
                   ggtitle("Similarity Heat Map") + labs(subtitle="Euclidean distance on variance stabilized data"),
                   rel_widths = c(1,.1,1), ncol=3)
dev.off()
cowplot::plot_grid(ggPCA + xlab("PC1: 97% variance") + ylab("PC2: 1% variance"),
                   ggplot() + theme_void(), 
                   hm + theme(plot.title = element_text(hjust = 0.5, face="bold"),
                                                       plot.subtitle = element_text(hjust = 0.5)) + 
                   ggtitle("Similarity Heat Map") + labs(subtitle="Euclidean distance on variance stabilized data"),
                   rel_widths = c(1,.1,1), ncol=3)
```
### RNA-Seq (Raw)
Sample TGFb_24h_R1 was sequenced deeper separated from the other samples. Therefore a lot of genes/transcripts are detected, which are not detected in the other samples. Consequently, it is popping up as an "outlier" in the similarity heatmap. In the PCA plot, you could see that the PC2 reflects the treatment effect, while the depth of sequencing/batch effect is mapped on the PC1. 

```{r fig.width=13, fig.height=6, echo=F}
pdf(file="Data/Images/QC.RNASeq.pdf", width=13, height=6)
cowplot::plot_grid(gex.ggPCA[[1]],
                                        ggplot() + theme_void(), 
                                        gex.hm[[1]] + theme( plot.title = element_text(hjust = 0.5, face="bold"), 
                                                        plot.subtitle = element_text(hjust = 0.5)) + 
                                                        ggtitle("Similarity Heat Map") + 
                                                        labs(subtitle="Euclidean distance on variance stabilized data"),
                                        rel_widths = c(1,.1,1),
                                        ncol=3)
dev.off()
cowplot::plot_grid(gex.ggPCA[[1]],
                                        ggplot() + theme_void(), 
                                        gex.hm[[1]] + theme( plot.title = element_text(hjust = 0.5, face="bold"), 
                                                        plot.subtitle = element_text(hjust = 0.5)) + 
                                                        ggtitle("Similarity Heat Map") + 
                                                        labs(subtitle="Euclidean distance on variance stabilized data"),
                                        rel_widths = c(1,.1,1),
                                        ncol=3)
```

### RNA-Seq (Batch correction)

After performing some batch correction, you could see that the depth of sequencing effect and the corresponding batch effect is largely removed. It seems, however, as if the batch correction is not specific enough, which is reflected by the similarity heatmap. However, on PC1, we now have a separation between the conditions. 

```{r fig.width=13, fig.height=6, echo=F}
pdf(file="Data/Images/QC.RNASeq.BatchCorrection.pdf", width=13, height=6)
cowplot::plot_grid(  gex.ggPCA[[2]],
                     ggplot() + theme_void(), 
                     gex.hm[[2]] + theme( plot.title = element_text(hjust = 0.5, face="bold"), 
                                          plot.subtitle = element_text(hjust = 0.5)) + 
                                          ggtitle("Similarity Heat Map") + 
                                          labs(subtitle="Euclidean distance on variance stabilized data"),
                                          rel_widths = c(1,.1,1),
                                          ncol=3)
dev.off()
cowplot::plot_grid(  gex.ggPCA[[2]],
                     ggplot() + theme_void(), 
                     gex.hm[[2]] + theme( plot.title = element_text(hjust = 0.5, face="bold"), 
                                          plot.subtitle = element_text(hjust = 0.5)) + 
                                          ggtitle("Similarity Heat Map") + 
                                          labs(subtitle="Euclidean distance on variance stabilized data"),
                                          rel_widths = c(1,.1,1),
                                          ncol=3)
```

## {.tabset .tabset-fade }

<h3>Correlation Plots</h3>

```{r echo=F, fig.height=7, fig.width=21}
corr_plot <- vector("list", 3)

gex.vsd <- as.data.frame(assay(rlog(DDS.GEX, fitType = "local", blind=F)))
gex.vsd$var <- matrixStats::rowVars(as.matrix(gex.vsd))
gex.vsd <- gex.vsd[order(gex.vsd$var),]

selected.genes_ids <- rownames(gex.vsd[1:50,])

atac.vsd <- as.data.frame(assay(rlog(DDS.atac, fitType = "local", blind=F)))
atac.vsd$var <- matrixStats::rowVars(as.matrix(atac.vsd))
atac.vsd <- atac.vsd[order(atac.vsd$var),]

selected.genes_ids.atac <- rownames(atac.vsd[1:50,])
atac.vsd$var <- NULL
gex.vsd$var <- NULL
atac.vsd$atac_avg_0h <- (atac.vsd[,1]+atac.vsd[,2])/2
atac.vsd$atac_avg_24h <- (atac.vsd[,3]+atac.vsd[,4])/2
atac.vsd$atac_avg_72h <- (atac.vsd[,5]+atac.vsd[,6])/2
gex.vsd$gex_avg_0h <- (gex.vsd[,1]+gex.vsd[,2])/2
gex.vsd$gex_avg_24h <- (gex.vsd[,3]+gex.vsd[,4])/2
gex.vsd$gex_avg_72h <- (gex.vsd[,5]+gex.vsd[,6])/2
gex.vsd <- gex.vsd[,7:9]
gex.vsd$input <- rownames(gex.vsd)
atac.vsd <- atac.vsd[,7:9]
atac.vsd$input <- rownames(atac.vsd)
colnames(GENES[[1]]) <- c("Symbol", "input", "ensembl_gene_id")
gex.vsd <- merge(gex.vsd, GENES[[1]], by="input")
gex.vsd <- gex.vsd[,c(5,6,2:4)]
atac.vsd <- merge(atac.vsd, GENES[[1]], by="input")
atac.vsd <- atac.vsd[,c(5,6,2:4)]

correlation <- merge(gex.vsd, atac.vsd, by=c("Symbol","ensembl_gene_id"))
unloadNamespace("ggpubr")
library(ggsignif)
corr_plot[[1]] <- ggstatsplot::ggscatterstats(
                                              data = correlation,
                                              x = atac_avg_0h,
                                              y = gex_avg_0h,
                                              xlab = "ATAC - regularized logarithm of counts (0h)",
                                              ylab = "GEX - regularized logarithm of counts (0h)",
                                              title = "Correlation plot (0h)"
                                            )
corr_plot[[2]] <- ggstatsplot::ggscatterstats(
                                              data = correlation,
                                              x = atac_avg_24h,
                                              y = gex_avg_24h,
                                              xlab = "ATAC - regularized logarithm of counts (24h)",
                                              ylab = "GEX - regularized logarithm of counts (24h)",
                                              title = "Correlation plot (24h)"
                                            )
corr_plot[[3]] <- ggstatsplot::ggscatterstats(
                                              data = (correlation),
                                              x = atac_avg_72h,
                                              y = gex_avg_72h,
                                              xlab = "ATAC - regularized logarithm of counts (72h)",
                                              ylab = "GEX - regularized logarithm of counts (72h)",
                                              title = "Correlation plot (72h)"
                                            )

pdf(file = "Data/Images/corrplots.pdf", width=21, height=7)
cowplot::plot_grid(corr_plot[[1]],corr_plot[[2]],corr_plot[[3]],ncol=3)
dev.off()
cowplot::plot_grid(corr_plot[[1]],corr_plot[[2]],corr_plot[[3]],ncol=3)

```

Here we show sample-wise correlation plots between GEX and ATAC information. On the y axis, we plot the average expression of each gene across replicates; on the x-axis, the average reads coverage of a gene around the TSS region (+/- 1KB). The blue line shows the fitting line. The correlation coefficient is computed using Pearson's method. The bar plot on the top side shows the distribution of the read coverage levels in the TSS region (+/- 1KB). The bar plot on the right side shows the distribution of the gene expression levels. Jointly the bar plots show that high gene expression generally corresponds to a high level of openness in chromatin. This fact is also reflected in Pearson's correlation coefficients.

## {.tabset .tabset-fade }

<h3>Heatmaps (Gene-Level)</h3>

```{r echo=F}
## Genes 
GENES <- vector("list", length(pairs))
Genes <- read.table("~/Desktop/work/RProjects/Data/Genes.HomoSapiens.gtf", sep="\t")
Genes <- Genes[,9]
Genes <- stringr::str_split(Genes, ";")
Gene.Names <- vector("list", length(Genes))
Gene.Ids <- vector("list", length(Genes))
for(g in 1:length(Gene.Names)){
  Gene.Names[[g]] <- stringr::str_trim(Genes[[g]][3])
  Gene.Ids[[g]] <- stringr::str_trim(Genes[[g]][1])
}
Gene.Names <- unlist(Gene.Names)
Gene.Names <- gsub(Gene.Names, pattern = "gene_name ", replacement = "")
Gene.Ids <- unlist(Gene.Ids)
Gene.Ids <- gsub(Gene.Ids, pattern = "gene_id ", replacement = "")
names(Gene.Names) <- Gene.Ids
```


### Top 50 variable genes heatmap 
```{r fig.width=16, fig.height=10, echo=F}


## Normalize count data
gex.vsd <- as.data.frame(assay(rlog(DDS.GEX, fitType = "local", blind=F)))
atac.vsd <- as.data.frame(assay(rlog(DDS.atac, fitType = "local", blind=F)))

### Gene Selection.
selected_genes <- vector("list",4)
for(i in 1:3){
   selected_genes[[i]] <-  DDS.GEX.results[[i]]
   selected_genes[[i]] <-  selected_genes[[i]][ selected_genes[[i]]$padj <= 0.05 &  selected_genes[[i]]$log2FoldChange >= 1,]
   selected_genes[[i]] <-  selected_genes[[i]][order( selected_genes[[i]]$padj),]
   selected_genes[[i]] <-  selected_genes[[i]][1:15,]
   selected_genes[[i]] <-  selected_genes[[i]]$`Symbol (HGNC)`
}
i <- 4
selected_genes[[i]] <- DDS.GEX.results[[i-1]]
selected_genes[[i]] <-  selected_genes[[i]][ selected_genes[[i]]$padj <= 0.05 &  selected_genes[[i]]$log2FoldChange <= -1,]
selected_genes[[i]] <-  selected_genes[[i]][order( selected_genes[[i]]$padj),]
selected_genes[[i]] <-  selected_genes[[i]][1:15,]
selected_genes[[i]] <-  selected_genes[[i]]$`Symbol (HGNC)`
selected_genes <- unique(unlist(selected_genes))

### Heatmap generation.
assays <- vector("list", 2)
assays[[1]] <- gex.vsd
assays[[2]] <- atac.vsd
heatmaps <- vector("list", length(assays))
legends <- c(F,T)
assay_titles <- c("GEX", "ATAC")
hjust <- c(.47,.3)
hm.gex.matrix <- vector("list", length(assays))
for(k in 1:length(assays)){
   GENES[[p]] <- Gene.Names #gprofiler2::gconvert(query = rownames(DDS.GEX.results[[p]]), organism = organism, target = target)
   GENES[[p]] <- as.data.frame(GENES[[p]])
   GENES[[p]]$input <- rownames(GENES[[p]])
   colnames(GENES[[p]]) <- c("Symbol", "input")
   gene_matrix <- as.data.frame(assays[[k]])
   hm.gex.matrix[[k]] <- gene_matrix
   hm.gex.matrix[[k]]$input <-rownames(hm.gex.matrix[[k]])
   hm.gex.matrix[[k]] <- merge(hm.gex.matrix[[k]], GENES[[p]], by="input")
   hm.gex.matrix[[k]] <- hm.gex.matrix[[k]][hm.gex.matrix[[k]]$Symbol %in% selected_genes, ]
   rownames(hm.gex.matrix[[k]]) <- hm.gex.matrix[[k]]$Symbol
   hm.gex.matrix[[k]] <- hm.gex.matrix[[k]][, 2:7]
   hm.gex.matrix[[k]] <- as.data.frame((hm.gex.matrix[[k]]))
   
   annotation_col <- data.frame(SampleType = factor(c("TGFb_0h","TGFb_0h","TGFb_24h","TGFb_24h","TGFb_72h","TGFb_72h")),
                                Treated    = factor(c("N", "N", "Y", "Y", "Y", "Y")))
   rownames(annotation_col) <- colnames(as.data.frame(hm.gex.matrix[[k]]))
   ann_colors <- list(SampleType = c(`TGFb_0h` = "darkgreen", `TGFb_24h`="red", `TGFb_72h`="orange"),
                      Treated = c(N="lightgrey", Y="green"))
   heatmaps[[k]] <- pheatmap::pheatmap(as.matrix(hm.gex.matrix[[k]]), main = "",
                                       color = colorRampPalette(c("darkblue", "white", "red2"))(50), 
                                       gaps_col = c(2,4),  border_color = "lightgrey", 
                                       annotation_colors = ann_colors, cluster_rows=ifelse(k==1, T,F), 
                                       annotation_col = annotation_col, show_rownames=T, show_colnames = F, 
                                       cluster_cols=F, fontsize_row = 10, scale = "row", 
                                       annotation_legend = legends[k], 
                                       legend=legends[k]) 
}

library(dplyr)
hm.gex.matrix[[1]]$id <- 1:nrow(hm.gex.matrix[[1]])
x <- heatmaps[[1]]$tree_row[["order"]]
hm.gex.matrix[[1]] <- hm.gex.matrix[[1]] %>% slice(match(x, id))
hm.gex.matrix[[1]]$Symbol <- rownames(hm.gex.matrix[[1]])
hm.gex.matrix[[2]]$Symbol <- rownames(hm.gex.matrix[[2]])
hm.gex.matrix[[2]] <- merge(hm.gex.matrix[[1]], hm.gex.matrix[[2]], by="Symbol", all.x=T)
hm.gex.matrix[[2]][is.na(hm.gex.matrix[[2]])] <- 0.01
hm.gex.matrix[[2]] <- hm.gex.matrix[[2]][, c(1,8:14)]
hm.gex.matrix[[2]] <- hm.gex.matrix[[2]] %>% slice(match(x, id))
rownames(hm.gex.matrix[[2]]) <- hm.gex.matrix[[2]]$Symbol
hm.gex.matrix[[1]]$Symbol <- NULL
hm.gex.matrix[[1]]$id <- NULL
hm.gex.matrix[[2]]$Symbol <- NULL
hm.gex.matrix[[2]]$id <- NULL

hjust <- c(.49,.325)
for(k in 1:length(assays)){
   annotation_col <- data.frame(SampleType = factor(c("TGFb_0h","TGFb_0h","TGFb_24h","TGFb_24h","TGFb_72h","TGFb_72h")),
                                Treated = factor(c("N", "N", "Y", "Y", "Y", "Y")))
   rownames(annotation_col) <- colnames(as.data.frame(hm.gex.matrix[[k]]))
   ann_colors <-list(SampleType = c(`TGFb_0h` = "#003f5c", `TGFb_24h`="purple", `TGFb_72h`="#ffa600"),
                      Treated = c(N="red", Y="green"))
   heatmaps[[k]] <- pheatmap::pheatmap(as.matrix(hm.gex.matrix[[k]]), main = "", na_col="grey", 
                                       color = colorRampPalette(c("darkblue", "white", "red2"))(50), 
                                       gaps_col = c(2,4),  border_color = "lightgrey", 
                                       annotation_colors = ann_colors, cluster_rows=ifelse(k==1, T,F), 
                                       annotation_col = annotation_col, show_rownames=T, show_colnames = F, 
                                       cluster_cols=F, fontsize_row = 10, scale = "row", 
                                       annotation_legend = legends[k], 
                                       legend=legends[k]) 
  heatmaps[[k]] <-ggplotify::as.ggplot(heatmaps[[k]]) + ggtitle(assay_titles[k]) + theme(plot.title=element_text(hjust=hjust[k], face = "bold"))
}

top.gex.heatmaps <- plot_grid(heatmaps[[1]]+theme(legend.position="none"), heatmaps[[2]], ggplot()+theme_void(), rel_widths=c(1.35,1.6,.3), ncol=3)
```


```{r fig.width=16, fig.height=10, echo=F}
library(DESeq2)
library(cowplot)
library(ggplot2)

## Normalize count data
gex.vsd <- as.data.frame(assay(rlog(DDS.GEX, fitType = "local", blind=F)))
atac.vsd <- as.data.frame(assay(rlog(DDS.atac, fitType = "local", blind=F)))

### Gene Selection
DAG_GENES <- T
if(DAG_GENES){ 
   selected_genes <- vector("list",4)
   for(i in 1:3){
      selected_genes[[i]] <-  DDS.results[[i]]
      selected_genes[[i]] <-  selected_genes[[i]][ selected_genes[[i]]$padj <= 0.05 &  selected_genes[[i]]$log2FoldChange >= 1,]
      selected_genes[[i]] <-  selected_genes[[i]][order( selected_genes[[i]]$padj),]
      selected_genes[[i]] <-  selected_genes[[i]][1:15,]
      selected_genes[[i]] <-  selected_genes[[i]]$`Symbol (HGNC)`
   }
   i <- 4
   selected_genes[[i]] <-  DDS.results[[i-1]]
   selected_genes[[i]] <-  selected_genes[[i]][ selected_genes[[i]]$padj <= 0.05 &  selected_genes[[i]]$log2FoldChange <= -1,]
   selected_genes[[i]] <-  selected_genes[[i]][order( selected_genes[[i]]$padj),]
   selected_genes[[i]] <-  selected_genes[[i]][1:15,]
   selected_genes[[i]] <-  selected_genes[[i]]$`Symbol (HGNC)`
   selected_genes <- unique(unlist(selected_genes))
} else {
   
   
   
}





### Heatmap generation.
assays <- vector("list", 2)
assays[[1]] <- atac.vsd
assays[[2]] <- gex.vsd
heatmaps.atac <- vector("list", length(assays))
legends <- c(F,T)
assay_titles <- c("ATAC","GEX")
hm.atac.matrix <- vector("list", length(assays))
for(k in 1:length(assays)){
   GENES[[p]] <- Gene.Names 
   GENES[[p]] <- as.data.frame(GENES[[p]])
   GENES[[p]]$input <- rownames(GENES[[p]])
   colnames(GENES[[p]]) <- c("Symbol", "input")
   gene_matrix <- as.data.frame(assays[[k]])
   hm.atac.matrix[[k]] <- gene_matrix
   hm.atac.matrix[[k]]$input <-rownames(hm.atac.matrix[[k]])
   hm.atac.matrix[[k]] <- merge(hm.atac.matrix[[k]], GENES[[p]], by="input")
   hm.atac.matrix[[k]] <- hm.atac.matrix[[k]][hm.atac.matrix[[k]]$Symbol %in% selected_genes, ]
   rownames(hm.atac.matrix[[k]]) <- hm.atac.matrix[[k]]$Symbol
   hm.atac.matrix[[k]] <- hm.atac.matrix[[k]][, 2:7]
   hm.atac.matrix[[k]] <- as.data.frame((hm.atac.matrix[[k]]))
   
   annotation_col <- data.frame(SampleType = factor(c("TGFb_0h","TGFb_0h","TGFb_24h","TGFb_24h","TGFb_72h","TGFb_72h")),
                               Treated = factor(c("N", "N", "Y", "Y", "Y", "Y")))
   rownames(annotation_col) <- colnames(as.data.frame(hm.atac.matrix[[k]]))
   ann_colors <- list(SampleType = c(`TGFb_0h` = "darkgreen", `TGFb_24h`="red", `TGFb_72h`="orange"),
                     Treated = c(N="lightgrey", Y="green"))
   heatmaps.atac[[k]] <- pheatmap::pheatmap(as.matrix(hm.atac.matrix[[k]]), main = "",
                                       color = colorRampPalette(c("darkblue", "white", "red2"))(50), 
                                       gaps_col = c(2,4),  border_color = "lightgrey", 
                                       annotation_colors = ann_colors, cluster_rows=ifelse(k==1, T, F), 
                                       annotation_col = annotation_col, show_rownames=T, show_colnames = F, 
                                       cluster_cols=F, fontsize_row = 10, scale = "row", 
                                       annotation_legend = legends[k], 
                                       legend=legends[k]) 
}

library(dplyr)


hm.atac.matrix[[1]]$id <- 1:nrow(hm.atac.matrix[[1]])
x <- heatmaps.atac[[1]]$tree_row[["order"]]
hm.atac.matrix[[1]] <- hm.atac.matrix[[1]] %>% slice(match(x, id))
hm.atac.matrix[[1]]$Symbol <- rownames(hm.atac.matrix[[1]])
hm.atac.matrix[[2]]$Symbol <- rownames(hm.atac.matrix[[2]])
hm.atac.matrix[[2]] <- merge(hm.atac.matrix[[1]], hm.atac.matrix[[2]], by="Symbol", all.x=T)
hm.atac.matrix[[2]][is.na(hm.atac.matrix[[2]])] <- 0.01
hm.atac.matrix[[2]] <- hm.atac.matrix[[2]][, c(1,8:14)]
hm.atac.matrix[[2]] <- hm.atac.matrix[[2]] %>% slice(match(x, id))
rownames(hm.atac.matrix[[2]]) <- hm.atac.matrix[[2]]$Symbol
hm.atac.matrix[[1]]$Symbol <- NULL
hm.atac.matrix[[1]]$id <- NULL
hm.atac.matrix[[2]]$Symbol <- NULL
hm.atac.matrix[[2]]$id <- NULL

hjust <- c(.48,.35)
for(k in 1:length(assays)){
   annotation_col <- data.frame(SampleType = factor(c("TGFB_0h","TGFB_0h","TGFB_24h","TGFB_24h","TGFB_72h","TGFB_72h")),
                                Treated = factor(c("N", "N", "Y", "Y", "Y", "Y")))
   rownames(annotation_col) <- colnames(as.data.frame(hm.atac.matrix[[k]]))
   ann_colors <- list(SampleType = c(`TGFB_0h` = "#003f5c", `TGFB_24h`="purple", `TGFB_72h`="#ffa600"),
                      Treated = c(N="red", Y="green"))
   heatmaps.atac[[k]] <- pheatmap::pheatmap(as.matrix(hm.atac.matrix[[k]]), main = "", na_col="grey", 
                                       color = colorRampPalette(c("darkblue", "white", "red2"))(50), 
                                       gaps_col = c(2,4),  border_color = "lightgrey", 
                                       annotation_colors = ann_colors, cluster_rows=ifelse(k==1, T,F), 
                                       annotation_col = annotation_col, show_rownames=T, show_colnames = F, 
                                       cluster_cols=F, fontsize_row = 10, scale = "row", 
                                       annotation_legend = legends[k], 
                                       legend=legends[k]) 
  heatmaps.atac[[k]] <-ggplotify::as.ggplot(heatmaps.atac[[k]]) + ggtitle(assay_titles[k])
}

top.atac.heatmaps <- plot_grid(heatmaps.atac[[1]]+theme(legend.position="none"), heatmaps.atac[[2]], ggplot()+theme_void(), rel_widths=c(1.35,1.,.1), ncol=3)
```

```{r fig.width=14, fig.height=9, echo=F}
library(cowplot)
library(ggplot2)

top.gex.heatmaps  <- plot_grid(heatmaps[[1]] + theme(plot.title = element_text(hjust = 0.48, face = "bold")) + theme(legend.position = "none"), 
                               heatmaps[[2]] + theme(plot.title = element_text(hjust = 0.12, face = "bold")), ggplot() + theme_void(), rel_widths = c(1.35, 1.9, .3), ncol=3)
title <- ggdraw() + draw_label("A. Top DEG Genes (RNA-Seq)", fontface = 'bold', x = 0, hjust = -0.5) + theme(plot.margin = margin(0, 0, 0, 7))
top.gex.heatmaps <- plot_grid(title, top.gex.heatmaps, ncol = 1, rel_heights = c(0.1, 1))

top.atac.heatmaps <- plot_grid(heatmaps.atac[[1]] + theme(plot.title = element_text(hjust = 0.46, face = "bold")) + theme(legend.position = "none"), 
                               heatmaps.atac[[2]] + theme(plot.title = element_text(hjust = 0.12, face = "bold")), ggplot() + theme_void(), rel_widths = c(1.35, 1.9, .1), ncol=3)
title <- ggdraw() + draw_label("B. Top DEA Genes (ATAC-Seq)", fontface = 'bold', x = 0, hjust = -0.5) + theme(plot.margin = margin(0, 0, 0, 7))
top.atac.heatmaps <- plot_grid(title, top.atac.heatmaps, ncol = 1, rel_heights = c(0.1, 1))

plot_grid(top.gex.heatmaps, plot_grid(top.atac.heatmaps, ggplot()+theme_void(), ncol=1, rel_heights = c(1,0.05)), rel_widths = c(1.05,1), ncol=2)

icesTAF::mkdir("Data/Images")
pdf(file="Data/Images/top50.heatmaps.pdf", width = 14, height = 9)
plot_grid(top.gex.heatmaps, plot_grid(top.atac.heatmaps, ggplot()+theme_void(), ncol=1, rel_heights = c(1,0.2)), rel_widths = c(1.05,1), ncol=2)
dev.off()
plot_grid(top.gex.heatmaps, plot_grid(top.atac.heatmaps, ggplot()+theme_void(), ncol=1, rel_heights = c(1,0.2)), rel_widths = c(1.05,1), ncol=2)

```

<h5>Caption: Two heatmaps pairs.</h5> 
<b>Left heatmap-pair:</b> after pooling the top 15 differentially expressed genes (DEG) of all comparisons (24h vs. 0h, 72 vs. 0h, 72 vs. 24h, and 24 vs. 72h), we perform hierarchical clustering of the genes based on their normalized expression values. The normalization is performed using the regularized logarithm as implemented by the DESeq2 package. Next, we present the results with a heatmap of the top genes with normalized expression values (left) and a corresponding heatmap showing the corresponding openness information from the ATAC-Seq assay. Grey cells correspond to genes with no detected ATAC-information. </br> 
<b>Right heatmap-pair:</b> after pooling the top 15 differentially accessible genes (DEA) of all comparisons (24h vs. 0h, 72 vs. 0h, 72 vs. 24h, and 24 vs. 72h), we perform hierarchical clustering of the genes based on their normalized accessibility values from the ATAC-Seq assay. The normalization is performed using the regularized logarithm as implemented by the DESeq2 package. Next, we present the results with a heatmap of the top genes with normalized accessibility values (left) and a corresponding heatmap showing the corresponding expression information from the RNA-Seq assay. Grey cells correspond to genes with no detected gene expression information. </br>

## {.tabset .tabset-fade }

<h3>Heatmaps (GeneSet-Level)</h3>

<h4>List of selected pathways</h4>
### GEX Selected Pathways 

```{r echo=F}
selected.gex.pathways <- c( "REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION",
                            "PID_INTEGRIN1_PATHWAY",
                            "REACTOME_COLLAGEN_FORMATION",
                            "REACTOME_COLLAGEN_BIOSYNTHESIS_AND_MODIFYING_ENZYMES",
                            "REACTOME_ELASTIC_FIBRE_FORMATION",
                            "PID_INTEGRIN3_PATHWAY",
                            "NABA_MATRISOME",
                            "REACTOME_INTERLEUKIN_4_AND_INTERLEUKIN_13_SIGNALING",
                            "REACTOME_INTEGRIN_CELL_SURFACE_INTERACTIONS",
                            "REACTOME_RUNX3_REGULATES_CDKN1A_TRANSCRIPTION",
                            "REACTOME_TRANSCRIPTIONAL_REGULATION_BY_RUNX3",
                            "NABA_COLLAGENS",
                            "REACTOME_RUNX3_REGULATES_NOTCH_SIGNALING",
                            "BIOCARTA_TGFB_PATHWAY",
                            "REACTOME_FIBRONECTIN_MATRIX_FORMATION",
                            "REACTOME_SIGNALING_BY_NOTCH1",
                            "REACTOME_SIGNALING_BY_PDGF",
                            "REACTOME_SIGNALING_BY_TGF_BETA_RECEPTOR_COMPLEX",
                            "NABA_ECM_REGULATORS",
                            "NABA_ECM_GLYCOPROTEINS")

selected.gex.pathways <- gsub(pattern = "REACTOME_", "", selected.gex.pathways)
selected.gex.pathways <- gsub(pattern = "PID_", "", selected.gex.pathways)
selected.gex.pathways <- gsub(pattern = "NABA_", "", selected.gex.pathways)
selected.gex.pathways <- gsub(pattern = "BIOCARTA_", "", selected.gex.pathways)
as.data.frame(selected.gex.pathways)
```

### ATAC Selected Pathways 

```{r echo=F}
selected.atac.pathways <- c("REACTOME_RUNX3_REGULATES_YAP1_MEDIATED_TRANSCRIPTION",
                            "REACTOME_SIGNALING_BY_NOTCH",
                            "NOTCH1_INTRACELLULAR_DOMAIN_REGULATES_TRANSCRIPTION",
                            "NOTCH_EXPRESSION_AND_PROCESSING",
                            "PID_PDGFRB_PATHWAY",
                            "SIGNALING_BY_TGF_BETA_RECEPTOR_COMPLEX",
                            "WNT_SIGNALING",
                            "MYC_PATHWAY",
                            "TRANSCRIPTIONAL_ACTIVITY_OF_SMAD2_SMAD3_SMAD4_HETEROTRIMER",
                            "SMAD2_SMAD3_SMAD4_HETEROTRIMER_REGULATES_TRANSCRIPTION",
                            "NABA_MATRISOME",
                            "NABA_MATRISOME_ASSOCIATED",
                            "REACTOME_TRANSCRIPTIONAL_REGULATION_BY_RUNX1",
                            "PID_PLK1_PATHWAY",
                            "REACTOME_SEMA4D_INDUCED_CELL_MIGRATION_AND_GROWTH_CONE_COLLAPSE",
                            "PID_PI3K_PLC_TRK_PATHWAY",
                            "PID_ERBB1_DOWNSTREAM_PATHWAY",
                            "PID_MAPK_TRK_PATHWAY",
                            "REACTOME_TNF_SIGNALING",
                            "REACTOME_SIGNALING_BY_TGF_BETA_FAMILY_MEMBERS")
selected.atac.pathways <- gsub(pattern = "REACTOME_", "", selected.atac.pathways)
selected.atac.pathways <- gsub(pattern = "PID_", "", selected.atac.pathways)
selected.atac.pathways <- gsub(pattern = "NABA_", "", selected.atac.pathways)
selected.atac.pathways <- gsub(pattern = "BIOCARTA_", "", selected.atac.pathways)
as.data.frame(selected.atac.pathways)
```

```{r echo=F}
go.tables <- vector("list", 4)
for(p in 1:4){
   if(p!=4){
      go.table <- rbind(gex.go_tbl_c2.DT[[p]]$x$data, gex.go_tbl_go.DT[[p]]$x$data)
   } else {
      go.table <- rbind(gex.go_Dn_tbl_c2.DT[[3]]$x$data, gex.go_Dn_tbl_go.DT[[3]]$x$data)
   }
go.table$GeneSetName <- gsub(pattern = "http://www.gsea-msigdb.org/gsea/msigdb/cards/", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "REACTOME_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "PID_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "NABA_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "BIOCARTA_", "", go.table$GeneSetName)
go.table <- go.table[go.table$GeneSetName %in% selected.gex.pathways,]
go.tables[[p]] <- go.table
}
```

## {.tabset .tabset-fade }

<h4>GEX: Geneset enrichment results of selected pathways</h4>
### Pathways in GOA (24h vs. 0h)
```{r echo=F}
DT::datatable(as.data.frame(go.tables[[1]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```


###  Pathways in GOA (72h vs. 0h)
```{r echo=F}
DT::datatable(as.data.frame(go.tables[[2]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```

### Pathways in GOA (72h vs. 24h)
```{r echo=F}
DT::datatable(as.data.frame(go.tables[[3]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```


```{r echo=F}
atac.go.tables <- vector("list", 3)
for(p in 1:4){
      
   if(p!=4){
      go.table <- rbind(go_tbl_c2.DT[[p]]$x$data, go_tbl_go.DT[[p]]$x$data)
   } else {
      go.table <- rbind(go_tbl_c2_dn.DT[[3]]$x$data, gex.go_Dn_tbl_go.DT[[3]]$x$data)
   }
go.table$GeneSetName <- gsub(pattern = "http://www.gsea-msigdb.org/gsea/msigdb/cards/", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "REACTOME_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "PID_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "NABA_", "", go.table$GeneSetName)
go.table$GeneSetName <- gsub(pattern = "BIOCARTA_", "", go.table$GeneSetName)
go.table <- go.table[go.table$GeneSetName %in% selected.atac.pathways,]
atac.go.tables[[p]] <- go.table
}
atac.go.tables


```


## {.tabset .tabset-fade }

<h4>ATAC: Geneset enrichment results of selected pathways</h4>
### Pathways in GOA (24h vs. 0h)
```{r echo=F}
DT::datatable(as.data.frame(atac.go.tables[[1]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```


###  Pathways in GOA (72h vs. 0h)
```{r echo=F}
DT::datatable(as.data.frame(atac.go.tables[[2]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```

### Pathways in GOA (72h vs. 24h)
```{r echo=F}
DT::datatable(as.data.frame(atac.go.tables[[3]]), rownames = FALSE, filter = 'top', 
                caption = Table_title, extensions = 'Buttons',
                options = list(dom = 'Bfrtip', scrollX=T,                                                                                                                          
                               buttons = list(list(extend='csv', title=table_name), 
                                              list(extend='excel', filename=table_name),    
                                              list(extend='pdf', filename=table_name)), 
                               autoWidth = TRUE, pageLength = 10, 
                               columnDefs = list(list(className = 'dt-left'))))
```

## {.tabset .tabset-fade }

<h4>GO heatmaps</h4>

### RNA-Seq
```{r echo=F}
colors <- c("#000080","#4c3197","#795eae","#a28dc4","#c9bedb",
               "#e3c4bc","#d29889","#bd6c5a","#a4402d","#880000",
               "#008000","#4c973c","#78ad68","#a1c495","#c9dac2",
               "#dcc5da","#c79ac3","#b170ac","#994396","#800080")

colors <- c("#ff8800","#ff9d45","#ffb271","#ffc79b","#fcdcc5",
               "#cadcc3","#a2c796","#7ab26a","#4e9d3e","#008800",
               "#58508d","#766ea1","#958db4","#b3adc8","#d2cedc",
               "#e8d1dd","#dfb2c9","#d492b6","#c972a3","#bc5090")
```

```{r echo=F}
  go.heatmap <- function(go.table, DEG.Table, cond, treat, gs_colors, vsd){
  

  
  ### Select genes from each pathway.
  DE <- vector("list", nrow(curr.go.table))
  lengths <- vector("list", nrow(curr.go.table))
  for(p in 1:nrow(curr.go.table)){
     DE[[p]] <- DEG.Table
     DE[[p]] <- DE[[p]][DE[[p]]$`Symbol (HGNC)` %in% unlist(strsplit(curr.go.table[p,]$Intersection,",")), ]
     DE[[p]] <- DE[[p]][order(DE[[p]]$padj),]
     DE[[p]] <- DE[[p]][1:5,]
     DE[[p]]$GeneClass <- ""
     DE[[p]] <- na.omit(DE[[p]])
     lengths[[p]] <- nrow(DE[[p]])
     DE[[p]]$GeneClass <- rep(curr.go.table[p,]$GeneSetName, lengths[[p]])
  }
  DE_all <- do.call("rbind", DE)
  DE_all <- DE_all[,c(1,2,ncol(DE_all))]

  ### Reduce count matrix to selected genes.
  gene_matrix <- vsd
  hm.gex.matrix <- gene_matrix
  hm.gex.matrix$ensembl_gene_id <-rownames(hm.gex.matrix)
  hm.gex.matrix <- merge(hm.gex.matrix, DE_all, by="ensembl_gene_id")
  rownames(hm.gex.matrix) <- 1:nrow(hm.gex.matrix)
  hm.gex.matrix <- hm.gex.matrix[,2:7]
  hm.gex.matrix <- as.data.frame((hm.gex.matrix))

  ### Annotate columns of the heatmap.
  annotation_col = data.frame(SampleType = factor(c(rep(cond[1],2), rep(cond[2],2), rep(cond[3],2))),
                              Treated = factor(c(rep(treat[1],2), rep(treat[2],2), rep(treat[3],2))))
  rownames(annotation_col) <- colnames(hm.gex.matrix)

  ### Annotate rows of the heatmap.
  annotation_row <- data.frame(GeneClass = factor(rep(curr.go.table$GeneSetName, unlist(lengths))))
  annotation_row$GeneClass <- factor(annotation_row$GeneClass, levels = unique(DE_all$GeneClass))
  rownames(annotation_row) = 1:nrow(annotation_row)

  ### select colors for row annotation
  
  names(gs_colors) <- curr.go.table$GeneSetName
  
  ann_colors = list(SampleType = c("darkgreen","red", "orange"),
                    Treated = c(N="lightgrey", Y="green"),
                    GeneClass=gs_colors
                    )
  names(ann_colors$SampleType) <- cond
  
  lengths <- unlist(lengths)
  gaps <- vector("list", length(lengths))
  gaps[[1]] <- lengths[1]
  for(l in 2:length(lengths)){ gaps[[l]] <- lengths[l] + gaps[[l-1]] }
  gaps <- unlist(gaps)

  names <- ifelse(endsWith(rownames(annotation_col), "R1"), "R1",
                  ifelse(endsWith(rownames(annotation_col), "R2"), "R2", "R3"))

  chm <- ComplexHeatmap::pheatmap(as.matrix(hm.gex.matrix), annotation_col = annotation_col, gaps_row = gaps, scale="row",
                                  annotation_row = annotation_row, annotation_colors = ann_colors,  legend = F,
                                  annotation_names_row = T, row_names_side="left", border_color = "lightgrey", labels_col = names,
                                  labels_row = DE_all$`Symbol (HGNC)`,   cluster_rows = F, cluster_cols = T,
                                  color = colorRampPalette(c("darkblue", "white", "red2"))(50))


  ha = HeatmapAnnotation(
      empty = anno_empty(border = FALSE, 
                         height = unit(4, "mm")),
                         foo = anno_block(gp = gpar(fill = 2:6),labels = c("0h", "24h", "72h"))
  )

  chm@left_annotation@anno_list$GeneClass@legend_param <- 
    list(labels_gp = gpar(fontsize = 10), plot = F, title = "C2 pathways", nrow = NULL,  ncol = 1,
         title_gp = gpar(fontsize = 10, fontface = "bold"), title_position = "topleft", color_bar = c("discrete"), 
         grid_height = unit(4, "mm"), grid_width = unit(4, "mm"), grid_border = NULL, labels_gp = gpar(fontsize = 10), 
         row_gap = unit(18, "mm"),  legend_height = NULL, legend_width = NULL, legend_direction = c("horizontal"))


# + ggtitle("RNA-Seq: [24h vs. 0h] - TOP GENES IN SELECTED PATHWAYS")

  gex.go.heatmap.1 <- ggplotify::as.ggplot(chm) + theme(plot.title = element_text(size=12, face = "bold", colour = "#000080", hjust=0.5))
return(gex.go.heatmap.1)
  }
```


```{r fig.width=14, fig.height=9, echo=F}
  pair <- 1
  cond <- c("TGFb_0h", "TGFb_24h", "TGFb_72h")
  treat <- c("N", "Y", "Y")
  DEG.Table <- DDS.GEX.results[[pair]]
  
  gs_colors <- colors[1:10]
  curr.go.table <- na.omit(go.tables[[pair]][1:10,])
  gex.go.heatmap.1 <- go.heatmap(go.table = curr.go.table, DEG.Table = DEG.Table, cond = cond, treat = treat, gs_colors=gs_colors, vsd=gex.vsd)
  gs_colors <- colors[11:20]
  curr.go.table <- na.omit(go.tables[[pair]][11:20,])
  gex.go.heatmap.2 <- go.heatmap(go.table = curr.go.table, DEG.Table = DEG.Table, cond = cond, treat = treat, gs_colors=gs_colors, vsd=gex.vsd)
  
  icesTAF::mkdir("Data/Images")
  pdf(file="Data/Images/gex.go.heatmap.pdf", width = 14, height = 9)
  cowplot::plot_grid(gex.go.heatmap.1, ggplot()+theme_void(), plot_grid(gex.go.heatmap.2, ggplot()+theme_void(), rel_heights = c(1,0.15), ncol=1),rel_widths=c(0.875,0.075,0.8), ncol=3)
  dev.off()
  cowplot::plot_grid(gex.go.heatmap.1, ggplot()+theme_void(), plot_grid(gex.go.heatmap.2, ggplot()+theme_void(), rel_heights = c(1,0.15), ncol=1),rel_widths=c(0.875,0.075,0.8), ncol=3)
```
### ATAC-Seq
```{r fig.width=17, fig.height=9, echo=F}
  pair <- 1
  cond <- c("TGFB_0h", "TGFB_24h", "TGFB_72h")
  treat <- c("N", "Y", "Y")
  DEG.Table <- DDS.results[[pair]]
  atac.go.table.all <- do.call("rbind",list(atac.go.tables[[1]],atac.go.tables[[3]]))
  
  
  gs_colors <- colors[1:10]
  curr.go.table <- na.omit(atac.go.table.all[1:10,])
  atac.go.heatmap.1 <- go.heatmap(go.table = curr.go.table, DEG.Table = DEG.Table, cond = cond, treat = treat, gs_colors=gs_colors, vsd=atac.vsd)
  gs_colors <- colors[11:19]
  curr.go.table <- na.omit(atac.go.table.all[(!atac.go.table.all$GeneSetName %in% atac.go.table.all[1:10,]$GeneSetName),])
  atac.go.heatmap.2 <- go.heatmap(go.table = curr.go.table, DEG.Table = DEG.Table, cond = cond, treat = treat, gs_colors=gs_colors, vsd=atac.vsd)
  
  
  icesTAF::mkdir("Data/Images")
  pdf(file="Data/Images/atac.go.heatmap.pdf", width = 17, height = 9)
  cowplot::plot_grid(atac.go.heatmap.1, ggplot()+theme_void(), plot_grid(atac.go.heatmap.2, ggplot()+theme_void(), rel_heights = c(1,0.15), ncol=1),rel_widths=c(0.875,0.075,0.8), ncol=3)
  dev.off()
  cowplot::plot_grid(atac.go.heatmap.1, ggplot()+theme_void(), plot_grid(atac.go.heatmap.2, ggplot()+theme_void(), rel_heights = c(1,0.15), ncol=1),rel_widths=c(0.875,0.075,0.8), ncol=3)

```

## {.tabset .tabset-fade }

<h3>Barplots (GeneSet-Level)</h3>

### RNA-Seq
```{r fig.width=16, fig.height=6, echo=F}
got <- vector("list", 3)
comparisons <- c("24h_vs_0h", "72h_vs_0h", "72h_vs_24h") 
for(i in 1:3){
   got[[i]] <- go.tables[[i]]
   got[[i]] <- got[[i]][,c("GeneSetName", "P_value")]
   got[[i]]$comparison <- comparisons[[i]]
}

got <- do.call("rbind", got)

got$comparison <- factor(got$comparison, levels = c("24h_vs_0h", "72h_vs_0h", "72h_vs_24h") )
gex.barplot <- ggplot(got, aes(x=GeneSetName, y=-log10(P_value), fill=comparison)) + geom_col() + coord_flip() + ggpubr::theme_pubr() + theme(axis.text.y = element_text(size=10)) + scale_fill_manual(values=c("darkgreen", "orange", "red")) + ylab("-log10(p-value) [Enrichment significance]") + facet_wrap(~comparison)

icesTAF::mkdir("Data/Images")
pdf(file="Data/Images/barplot.rnaseq.pdf", width = 16, height = 6)
gex.barplot
dev.off()
```

### ATAC-Seq

<h3>Barplots (GeneSet-Level)</h3>
```{r fig.width=16, fig.height=6, echo=F}
got <- vector("list", 3)
comparisons <- c("24h_vs_0h", "72h_vs_0h", "72h_vs_24h") 
for(i in 1:3){
   got[[i]] <- atac.go.tables[[i]]
   got[[i]] <- got[[i]][,c("GeneSetName", "P_value")]
   got[[i]]$comparison <- comparisons[[i]]
}

got <- do.call("rbind", got)
#got$log10pvalue <- ifelse(got$UP,-log10(got$P_value),log10(got$P_value))
#got$ExtGenesetName <- ifelse(got$UP,paste(got$GeneSetName," [UP]",sep=""),paste(got$GeneSetName," [DN]",sep=""))

got$comparison <- factor(got$comparison, levels = c("24h_vs_0h", "72h_vs_0h", "72h_vs_24h") )
atac.barplot <- ggplot(got, aes(x=GeneSetName, y=-log10(got$P_value), fill=comparison)) + geom_col() + coord_flip() + ggpubr::theme_pubr() + theme(axis.text.y = element_text(size=10)) + scale_fill_manual(values=c("darkgreen", "orange", "red")) + ylab("-log10(p-value) [Enrichment significance]") + facet_wrap(~comparison)
icesTAF::mkdir("Data/Images")

pdf(file="Data/Images/barplot.atacseq.pdf", width = 16, height = 6)
atac.barplot
dev.off()
```

## {.tabset .tabset-fade }

### SessionInfo
```{r echo=F}
sessionInfo()
```


